### 2022-11-17 chapter1/biorxiv.tex
Added biorxiv.tex

\documentclass[11pt,letterpaper]{article}

% PACKAGE CONFIGURATIONS
\usepackage{amsmath}

\usepackage{graphicx}

\usepackage[margin=1in]{geometry}
%\usepackage{setspace}  % double spacing
%\doublespacing

\usepackage[skip=10pt]{parskip}  % Remove indents

\usepackage{hyperref}

\usepackage[doi=false,url=false]{biblatex}
\addbibresource{../refs.bib}

\usepackage[labelfont=bf, labelsep=period]{caption}
\renewcommand{\figurename}{Fig}

% DOCUMENT
\begin{document}
\graphicspath{{./figures/}}

% Title must be 250 characters or less
\begin{flushleft}
{\Large\textbf{Leveraging genomic redundancy to improve inference and alignment of orthologous proteins}}
\newline
\\
Marc Singleton\textsuperscript{1},
Michael Eisen\textsuperscript{1,2,3,*}
\\
\bigskip
\textbf{1} Biophysics Graduate Group, UC Berkeley, Berkeley, CA, USA
\\
\textbf{2} Department of Molecular and Cell Biology, UC Berkeley, Berkeley, CA, USA
\\
\textbf{3} Howard Hughes Medical Institute, UC Berkeley, Berkeley, CA, USA
\\
\bigskip

* mbeisen@berkeley.edu

\end{flushleft}

\input{chapter1.tex}

\printbibliography

\clearpage

\input{chapter1_supporting.tex}

\end{document}