### 2023-02-15 chapter1/biorxiv.tex
Removed section numbering from chapter1.tex and reorganized imports

\documentclass[11pt,letterpaper]{article}

% PACKAGE CONFIGURATIONS
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[labelfont=bf, labelsep=period]{caption}
\renewcommand{\figurename}{Fig}

\usepackage[margin=1in]{geometry}
%\usepackage{setspace}  % double spacing
%\doublespacing
\usepackage[skip=10pt]{parskip}  % Remove indents
\setcounter{secnumdepth}{0}  % Remove section numbering

\usepackage{hyperref}
\usepackage[doi=false,url=false]{biblatex}
\addbibresource{../refs.bib}

% DOCUMENT
\begin{document}
\graphicspath{{./figures/}}

% Title must be 250 characters or less
\begin{flushleft}
{\Large\textbf{Leveraging genomic redundancy to improve inference and alignment of orthologous proteins}}
\\
\bigskip
Marc Singleton\textsuperscript{1,*},
Michael Eisen\textsuperscript{1,2,3}
\\
\bigskip
\textbf{1} Biophysics Graduate Group, UC Berkeley, Berkeley, CA, USA
\\
\textbf{2} Department of Molecular and Cell Biology, UC Berkeley, Berkeley, CA, USA
\\
\textbf{3} Howard Hughes Medical Institute, UC Berkeley, Berkeley, CA, USA
\\
\bigskip

* marcsingleton@berkeley.edu

\end{flushleft}

\input{chapter1.tex}

\section{Acknowledgments}
We thank the past and present members of the Eisen lab for helpful discussions and their assistance in revising the manuscript.

\printbibliography

\clearpage

\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}

\section*{Supporting information}
\input{chapter1_supporting.tex}

\end{document}