### 2023-02-22 main.tex
Re-named chapter 3

\documentclass[11pt,letterpaper,oneside]{book}

% PACKAGE AND ENVIRONMENT CONFIGURATIONS
\usepackage{amsmath}
\usepackage{graphicx}

% Code listings
\usepackage[outputdir=out]{minted}
\usepackage{tcolorbox}
\usepackage{lmodern}

% Margins and spacing
\usepackage[margin=1in]{geometry}
%\usepackage{setspace}  % double spacing
%\doublespacing
\usepackage[skip=10pt]{parskip}  % Remove indents
\pagestyle{plain}

% Chapter and figure formatting
\usepackage{titlesec}  % Modify chapter title format
\titleformat{\chapter}[display]% Shape (chapter label in separate paragraph)
                      {\bfseries\Large}% Format
                      {\MakeUppercase{\chaptertitlename} \thechapter}% Label
                      {0in}% Sep
                      {}% Before code
                      []% After code
\setcounter{tocdepth}{1}  % Only number divisions to section
\usepackage[labelfont=bf, labelsep=period]{caption}
\renewcommand{\figurename}{Fig}

% References
\usepackage[doi=false,url=false]{biblatex}
\addbibresource{refs.bib}
\usepackage{hyperref}

% ENVIRONMENTS
\newenvironment{abstract}
{% Begin code
 \small
 \begin{center}
 \textbf{Abstract}
 \end{center}
 \quotation
}
{% End code
}

\definecolor{incolor}{RGB}{51,63,153}
\definecolor{outcolor}{RGB}{200,78,41}
\definecolor{bgcolor}{RGB}{247,247,247}

\newlength{\promptwidth}
\setlength{\promptwidth}{0.5in}

\newcommand{\prompt}[2]
{\makebox[0pt][r]{\texttt{\color{#2}#1:\ }}\vspace{-\baselineskip}}

\newenvironment
{NotebookIn}
{
 \VerbatimEnvironment
 \begin{tcolorbox}[sharp corners,
                   frame empty,
                   size=fbox,
                   colback=bgcolor,
                   left skip=\promptwidth,
                   title=\prompt{In}{incolor},
                   attach title to upper]
 \begin{minted}{python}
}
{
 \end{minted}
 \end{tcolorbox}
}

\newenvironment
{NotebookOut}
{
 \begin{tcolorbox}[sharp corners,
                   frame empty,
                   size=fbox,
                   colback=white,
                   left skip=\promptwidth,
                   title=\prompt{Out}{outcolor},
                   attach title to upper]
 \\ \ttfamily
}
{
 \end{tcolorbox}
}

\newenvironment
{NotebookImage}
{
 \begin{tcolorbox}[sharp corners,
                   frame empty,
                   size=fbox,
                   colback=white,
                   left skip=\promptwidth,
                   title=\prompt{Out}{outcolor},
                   attach title to upper]
 \\
}
{
 \end{tcolorbox}
}

% DOCUMENT
\begin{document}

\begin{titlepage}
\begin{center}
    My long and grand sounding dissertation title\\
    \bigskip
    By\\
    Marc Singleton\\
    \vfill
    A dissertation submitted in partial satisfaction of the\\
    requirements for the degree of\\
    Doctor of Philosophy\\
    in\\
    Biophysics\\
    in the\\
    Graduate Division\\
    of the\\
    University of California, Berkeley\\
    \vfill
    Committee in charge:\\
    Professor Mike Eisen, Chair\\
    Professor Liana Lareau\\
    Professor Susan Marqusee\\
    Professor Priya Moorjani\\
    \bigskip
    Spring 2023\\
\end{center}
\end{titlepage}

\thispagestyle{empty}
\null  % Necessary to have text or command before \vfill for proper spacing
\vfill
\begin{center}
    \copyright \  2023 Marc Singleton
\end{center}
\clearpage

\setcounter{page}{1}  % Reset page counter after copyright
\begin{center}
    Abstract\\
    \bigskip
    My long and grand sounding dissertation title\\
    by\\
    Marc Singleton\\
    \bigskip
    Doctor of Philosophy in Biophysics\\
    University of California, Berkeley\\
    Professor Mike Eisen, Chair\\
    \bigskip
\end{center}
Here's the text for my abstract!
\clearpage

\setcounter{page}{1}  % Reset page counter after abstract
\pagenumbering{roman}  % Set numbering style to roman for pre-main content

\null  % Necessary to have text or command before \vfill for proper spacing
\vfill
\begin{center}
    To my parents, Amy and David, who were always willing\\
    to indulge another ``experiment'' in their kitchen sink
\end{center}
\vfill
\clearpage

\begin{center}
    \bfseries\Large Acknowledgements
\end{center}
\bigskip
If it takes a village to raise a child, then it must take at least three to raise a scientist. Over the years, I've had countless friends, teachers, and mentors. These relationships have shaped the person I am today, so I'm grateful to all the people who took time out of their busy lives to help me in ways both small and large. However, I want to single out the following people who had an especially important impact on my time in graduate school.\\\\
Thank you to my advisor, Mike, for supporting me as a scientist and a person in equal measure.\\\\
Thank you to my committee, Susan, Liana, and Priya, for giving me the pep talk I needed every year.\\\\
Thank you to my lab, Holli, Stadler, Ciera, Ashley, Victoria, Augusto, and Xiao-Yong, for creating a space where it's okay to ask questions. A special thank you to my bay mates, Jenna and Colleen, for getting me through the hard days with sound advice and a healthy dose of humor.\\\\
Thank you to my undergraduate research mentors, Cathy and Jessica, for teaching me the most exciting science is at the intersection of different fields.\\\\
Thank you to my parents, for everything.\\\\
Thank you to my brothers, Andrew and Daniel, for giving me thick skin in the way only big brothers can.\\\\
Thank you to all the friends I've made in California and the Bay Area, for the much-needed life outside work.\\\\
Finally, thank you to my partner, Joseph, for going on this crazy ride with me. I couldn't have done this without you.
\clearpage

\tableofcontents

\setcounter{chapter}{-1}  % Set chapter counter to 0
\chapter{Introduction}
\graphicspath{{chapter0/figures/}}
\setcounter{page}{1}  % Reset page counter after pre-main content
\pagenumbering{arabic}  % Set numbering style to arabic for main content
\input{chapter0/chapter0.tex}

\chapter{Leveraging genomic redundancy to improve inference and alignment of orthologous proteins}
\graphicspath{{chapter1/figures/}}
\input{chapter1/chapter1.tex}

\chapter{Evolutionary analyses of IDRs reveal conserved features}
\input{chapter2/chapter2.tex}

\chapter{Tools and tutorials for fitting mixture models and HMMs}
\graphicspath{{chapter3/}}
\input{chapter3/introduction.tex}

\printbibliography[heading=bibintoc]

\appendix
\setcounter{figure}{0}
\renewcommand{\thefigure}{\thechapter\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{\thechapter\arabic{table}}

\chapter{Supporting information for chapter 1}
\graphicspath{{chapter1/figures/}}
\input{chapter1/chapter1_supporting.tex}

\end{document}